---
title: "eDNAPlus vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{eDNAPlus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r, include = FALSE}
library(eDNAPlus)
```


The fitting is performed using the function `runEDNA`. The functions fits the model described in Diana et al. (2022). The user can choose to fit a model accounting for species correlation using the parameter `jointSpecies`.

## Data format

Sample data to feed to the function `runEDNA` are provided in the object `sampleData`. 


The data are encoded in a list formed by the objects:

* PCR_table

```{r}
head(data$PCR_table, 3)
```

Data.frame of species reads with environmental samples on the rows and species on the columns. Each PCR is formed by a group of columns. For example, if there are 3 PCR and 10 species, columns 1-10 corresponds to the first PCR, columns 11-20 to the second PCR, and so on. The species names are specified in the first set of column names.

* PCR_spike

Data.frame of spike-in species reads. Structure analogous to the previous.


* infos

```{r}
head(data$infos, 7)
```

A data.frame containing the information of the samples in the PCR table. The columns are:

   1. Site: Site from which the sample was taken. If it is an empty tube, site should be "empty"

   2. Samples: Name of the samples.

   3. PCR: Number of PCR replicates in this sample.


* X_z: Biomass abundance covariates matrix

```{r}
head(data$X_z, 7)
```

A matrix with sites on the rows and covariates on the columns. The name of the site is in the first column. To match with the `infos` object, the names of the sites in the two objects have to be the same.


* X_w: Biomass detection covariates matrix

```{r}
head(data$X_w, 7)
```

A matrix with samples on the rows and covariates on the columns. The name of the sample is in the first column. To match with the `infos` object, the names of the samples in the two objects have to be the same.


## Priors

Even if not necessary, the user can choose their own prior. The prior parameters that can be chosen are:

* `sigma_u`: Variance of the PCR bias `u`.

* `a_sigma`, `b_sigma`: Parameters of the inverse-gamma prior on the samples variance `sigma`.

For further explanation of each parameter, the user can refer to the paper.

# Model fitting

```{r, results = "hide"}
#modelResults <- runEDNA(data,
 #                       priors,
  #                      jointSpecies = F,
   #                     MCMCparams = MCMCparams) 
```



## Analyzing the output

The model output can be analyzed through the functions provided with the package. We provide functions for plotting summaries and printing outputs.

For plotting, we provide the functions:

* `plotOccupancyIndex`: Plots estimates of the occupancy index.

```{r,  fig.width = 6, fig.height= 4}
#plotOccupancyIndex(modelResults)
```

* `plotSpatialSiteEffect`: Plots a map with the posterior means of the spatial site effect.

## Diagnostics

The package provides function to produce traceplots to assess convergence. The functions provided are:

* `tracePlot_OccupancyYearEffect`: Traceplots of the year-specific random effects.

```{r, fig.width = 6, fig.height= 4}
#tracePlot_OccupancyYearEffect(modelResults, 1)
```

* `tracePlot_OccupancySiteEffect`: Traceplots of the site-specific random effects.

* `tracePlot_OccupancyCovariate`: Traceplot of the covariates of the occupancy probability.

* `tracePlot_DetectionIntercept`: Traceplot of the intercept of the detection probability.

* `tracePlot_DetectionCovariate`: Traceplot of the covariates of the detection probability.
